pema application

### 2022-12-07 paper/action_letter.Rmd
update 2022-12-07 15:57:04

We recognize that this choice of words was confusing.
The regularized horseshoe prior is an extension of the horseshoe prior.
To address this comment, we now explain this in the text:

@@

We agree with the Reviewer that there are important limitations to meta-regression, and Thompson & Higgins provide an excellent analysis thereof.
<!-- Limitations -->
<!-- i) multicolinearity -->
<!-- ii) ecological fallacy -->
<!-- iii) restriction of range -->
<!-- iv) multicolinearity -->
<!-- v) something specific to clinical trials?? -->
<!-- vi) data availability -->
<!-- vii) small sample size, limited generalizability -->
<!-- PITFALLS -->
<!-- i) curse of dimensionality: small sample size, multicolinearity -->
<!-- ii) overfitting -->
<!-- iii) variable selection -->
<!-- iv) variable selection -->
Most of the limitations and pitfalls they discuss come down to two fundamental points which are already addressed throughout our paper: 1) the curse of dimensionality and its implications for multicolinearity (limitations i, iv, vii and pitfalls i, ii, iii, and iv), and 2) the ecological fallacy (limitation ii and iv).
They further address three singular points which cannot be attributed to the two fundamental points:
Limitation iii relates to the problem of restriction of range in moderators (which we do not address),
Limitation vi relates to limited data availability (which we address in our discussion of missing data),
and Limitation v makes a point about regression to the mean in RCTs that we fail to grasp.

>>>>

We recognize that this choice of words was confusing.
The regularizing horseshoe prior is an extension of the horseshoe prior.
To address this comment, we now explain this in the text:

@@

We agree with the Reviewer that there are important limitations to meta-regression, and Thompson & Higgins provide an excellent analysis thereof.
<!-- Limitations -->
<!-- i) multicollinearity -->
<!-- ii) ecological fallacy -->
<!-- iii) restriction of range -->
<!-- iv) multicollinearity -->
<!-- v) something specific to clinical trials?? -->
<!-- vi) data availability -->
<!-- vii) small sample size, limited generalizability -->
<!-- PITFALLS -->
<!-- i) curse of dimensionality: small sample size, multicollinearity -->
<!-- ii) overfitting -->
<!-- iii) variable selection -->
<!-- iv) variable selection -->
Most of the limitations and pitfalls they discuss come down to two fundamental points which are already addressed throughout our paper: 1) the curse of dimensionality and its implications for multicollinearity (limitations i, iv, vii and pitfalls i, ii, iii, and iv), and 2) the ecological fallacy (limitation ii and iv).
They further address three singular points which cannot be attributed to the two fundamental points:
Limitation iii relates to the problem of restriction of range in moderators (which we do not address),
Limitation vi relates to limited data availability (which we address in our discussion of missing data),
and Limitation v makes a point about regression to the mean in RCTs that we fail to grasp.

### 2022-12-07 paper/manuscript.Rmd
update 2022-12-07 15:57:04

Meta-regression models are most commonly estimated using weighted least squares,
with weights determined according to the aforementioned random effects model,
where residual heterogeneity is estimated using restricted maximum likelihood [@pattersonRecoveryInterblockInformation1971; @viechtbauerBiasEfficiencyMetaAnalytic2005].
This approach, henceforth referred to as RMA, has low bias, which means that, across hypothetical replications of a study, the average values of model parameters are close to their true population values [@panityakulEstimatingResidualHeterogeneity2013].
Low bias comes at the cost of higher variance, however,
which means that the estimated values of population parameters vary more from one hypothetical replication to the next.
This phenomenon is known as the *bias-variance trade-off*.
In general, an estimator with low bias and high variance produces results that generalize worse to new data than an estimator with high bias and low variance.
Regularized approaches to regression intentionally increase bias in order to reduce variance.
This is a sensible approach in the context of small samples, which are common in meta-analyses,
because small samples incur a high risk of overfitting,
and typically have relatively high levels of multicollinearity,
due to the higher probability that extreme values on one moderator coincide with extreme values on another.
In such cases, regularized regression reduces the risk of overfitting and increase generalizability of the results [see @hastieElementsStatisticalLearning2009].

<span id = "reftibshirani">To understand how regularized regression introduces bias, consider a comparison between ordinary least squares regression and LASSO regression [for a more elaborate introduction, see @tibshiraniRegressionShrinkageSelection2011].</span>
Ordinary least squares regression (OLS) estimates model parameters by minimizing the Residual Sum of Squares (RSS) of the outcome variable, given by the formula below.
In this equation, $n$ is the number of participants; $y_i$ is the outcome variable, and $x$ is one of $p$ predictor variables.
\begin{align}
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
\end{align}
The resulting parameter estimates perfectly describe linear relations in the present data set, but do not generalize well to new data.
Regularized regression biases parameter estimates towards zero by adding a penalty term to the RSS.
Most common is the LASSO penalty, which consists of the sum of the absolute regression coefficients, or the L1 norm [@hastieElementsStatisticalLearning2009].
As the LASSO penalty is a function of the regression coefficients, it increases when they get bigger.
This incentivizes the optimizer to keep the regression coefficients as small as possible.
The amount of regularization can be controlled by multiplying the penalty by a tuning parameter, $\lambda$.
If $\lambda$ is zero, the shrinkage penalty has no impact.
As $\lambda$ increases,
all coefficients shrink towards zero, ultimately producing the null model.
Cross-validation is often used to find the optimal value for the penalty parameter $\lambda$.
The LASSO penalized residual sum of squares is given by:
\begin{align}
PRSS= RSS + \lambda \sum_{j=1}^{p}|\beta_{j}| (\#eq:prss)
\end{align}
<span id = "whylasso">Note that many other regularizing penalties exist.
This introduction focuses on the LASSO penalty because it is most ubiquitous, easy to understand,
and has an analogue in Bayesian estimation,
as explained in the next section.</span>

@@

A regularizing prior distribution shrinks small coefficients towards zero by assigning high probability mass to near-zero values.
There are many different regularizing prior distributions,
some of which are analogous to specific frequentist methods [@vanErpOberskiMulder2019].
For example, a double exponential prior (hereafter: LASSO prior) results in posterior distributions whose modes are identical to the estimates from LASSO-penalized regression [@ParkCasella2008].
Both the frequentist LASSO penalty and the Bayesian LASSO prior have a tuning parameter $\lambda$ that controls the amount of regularization.
In frequentist LASSO, its value is usually chosen via cross-validation [@hastieElementsStatisticalLearning2009]. 
In the Bayesian approach, by contrast,
its value can be estimated during model estimation.

@@

The BRMA method introduced here offers both LASSO and regularized horseshoe priors.
The LASSO prior is given by:
\begin{align}
\beta_{j} &\sim \text{DE}(0, \frac{s}{\lambda})\\ (\#eq:lasso)
\frac{1}{lambda} &\sim \chi^2(0, \nu_1)
\end{align}
where DE denotes the double exponential distribution centered around zero,
with a scale determined by a global scale parameter $s$ that is multiplied by the inverse of tuning parameter $\lambda$.
Increasing the scale parameter extends the prior to cover more extreme values.
The inverse tuning parameter is estimated from the data 
by assigning it a $\chi^2$ prior distribution with mean zero and degrees of freedom $\nu_1$.
Increasing the degrees of freedom assigns greater probability mass to extreme values, resulting in less regularization.
The present study used default values for the prior parameters,
$s = 1$, $\nu_1 = 1$, which are the same defaults as used in `brms` [@R-brms_a].

@@

We performed a simulation study to validate the BRMA algorithm.
As a benchmark for comparison, we used random effects meta-regression with restricted maximum likelihood estimation [RMA, @viechtbauerConductingMetaanalysesMetafor2010],
which is the current state-of-the-art in the field.
We evaluated the algorithms' predictive performance in new data, 
ability to perform variable selection,
and ability to recover population parameters.
Our research questions are whether BRMA offers a performance advantage over RMA in terms of any of these indicators,
and which prior (LASSO versus regularized horseshoe) is to be preferred.
<span id = "defaultvalues">For both Bayesian priors, we used default values proposed in prior literature, see Table \@ref(tab:tabprior).
Default values for the LASSO prior were based on @vanErpOberskiMulder2019,
and default values for the regularized horseshoe prior were based on @PiironenVehtari2017a.</span>
All analysis code is available in a version-controlled repository at <https://github.com/cjvanlissa/pema>.

@@

This example uses the `bonapersona` data, which were included in the `pema` package with permission of the author [@bonapersonaBehavioralPhenotypeEarly2019].
<span id = "sample400">This meta-analysis of over 400 experiments investigated the effects of early life adversity on cognitive performance in rodents.
Note that the sample is much larger than the maximum used to validate BRMA in our simulation study.
As larger samples provide greater statistical power,
it should also be valid for this sample.</span>
For illustrative purposes, we use a smaller subset of the more than 30 moderators.
See the `pema` package documentation (help and vignettes) for further examples.
<!-- We focus specifically on how to choose the different hyperparameters of the regularizing horseshoe prior. -->

```{r exdata, eval = FALSE, echo = TRUE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
```
```{r, eval = run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
saveRDS(df, "df_imputed.RData")
```
```{r, eval = !run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
df <- readRDS("df_imputed.RData")
```
<!-- Our simulation study shows good performance with default hyperparameters. -->
<!-- However, experienced users may want to customize the prior. -->
<!-- Visualizing the prior can be helpful in this process. -->
<!-- This is accomplished using the interactive application visualization application available through `shiny_prior()`. -->
<!-- The user can plot the prior distributions resulting from different sets of hyperparameters and compare them. -->
<!-- Increasing the values of the scale parameters (`scale_global` and `hs_scale_slab`) results in a more spread out prior, which applies less regularization. -->
<!-- Increasing the degrees of freedom (`df_global` and `df_slab`) results in thinner tails, which applies more regularization. -->
<!-- One advantage of the regularizing horseshoe prior is its ability to define the prior in terms of the expected number of nonzero coefficients, using the argument `relevant_pars`. -->
<!-- This argument in turn influences the other hyperparameters. -->
<!-- We can also plot the implied prior on the effective number of nonzero coefficients given specific hyperparameters. -->
<!-- In this application, we have a total of 7 possible moderators, including the dummy variables for the categorical moderator "model". We can visualize the implied prior on the number of nonzero coefficients for different values of `relevant_pars`: p0 = 1; p0 = 3; p0 = 6; or p0 = NA, in which case no prior information is given and `scale_global`= 1. -->

@@

```{r, echo = FALSE, eval = FALSE}
summary(fit)
```
```{r tabsumex, eval = TRUE, echo = FALSE, results = "asis"}
tab <- summary(fit)$coefficients[1:8, c("mean", "sd", "2.5%", "50%", "97.5%", "n_eff", "Rhat")]
papaja::apa_table(tab, caption = "Summary of model parameters for the applied example.")
```
<!-- , see Figure \@ref(fig:figexres) -->
```{r exres, eval = FALSE, echo = FALSE}
fit_stan <- as.stan(fit)
plot(fit_stan, pars = rownames(coef(fit))[2:8])
```
```{r, eval = run_everything, echo = FALSE}
fit_stan <- as.stan(fit)
p <- plot(fit_stan, pars = rownames(coef(fit))[2:8])
ggsave("plot_es.png", p, device = "png", width = 4, height = 3, dpi = 300, scale = 1.5)
```
```{r figexres, eval = FALSE, echo = FALSE, fig.cap="Visual presentation of model parameters for the applied example."}
knitr::include_graphics("plot_es.png")
```
<!-- Note that there are many auxiliary parameters used for the regularizing horseshoe prior. No information on these parameters is returned in the `brma` fitobject, however, once the fitobject is transformed into a Stan object for plotting, these parameters are included. -->

@@

<span id = "sensitivityanalysis">All simulations were conducted with default settings for the model's prior distributions, based on prior research [@vanErpOberskiMulder2019; @PiironenVehtari2017b].
Our results suggest that these defaults are suitable for a wide range of situations, including when model assumptions are violated.
However, bear in mind that model parameters are influenced by the prior distribution.
It is good practice to perform a sensitivity analysis to determine how sensitive the model results and inferences are to different prior specifications. 
Performing sensitivity analyses is particularly important when the sample is small,
as in this case, the prior is more influential.</span>

@@

<span id = "metareglimits">Before conducting meta-regression,
researchers should be aware of its limitations [see @thompsonHowShouldMetaregression2002].
These can be subdivided into four categories: 1) the curse of dimensionality and its corrolary implications for multicolinearity; 2) the ecological fallacy; 3) limited information on moderator variables, including missing data and restrictions of range.
BRMA seeks to address the first of these limitations,
because the problems that arise from meta-analyzing small and heterogeneous bodies of literature are so ubiquitous that they have been referred to as the primary pitfall in meta-regression [@thompsonHowShouldMetaregression2002].
Nonetheless, all applicable limitations should be acknowledged in the resulting publication.</span>

>>>>

Meta-regression models are most commonly estimated using weighted least squares,
with weights determined according to the aforementioned random effects model,
where residual heterogeneity is estimated using restricted maximum likelihood [@pattersonRecoveryInterblockInformation1971; @viechtbauerBiasEfficiencyMetaAnalytic2005].
This approach, henceforth referred to as RMA, has low bias, which means that, across hypothetical replications of a study, the average values of model parameters are close to their true population values [@panityakulEstimatingResidualHeterogeneity2013].
Low bias comes at the cost of higher variance, however,
which means that the estimated values of population parameters vary more from one hypothetical replication to the next.
This phenomenon is known as the *bias-variance trade-off*.
In general, an estimator with low bias and high variance produces results that generalize less well to new data than an estimator with high bias and low variance.
Regularized approaches to regression intentionally increase bias in order to reduce variance.
This is a sensible approach in the context of small samples, which are common in meta-analyses,
because small samples incur a high risk of overfitting,
and typically have relatively high levels of multicollinearity,
due to the higher probability that extreme values on one moderator coincide with extreme values on another.
In such cases, regularized regression reduces the risk of overfitting and increase generalizability of the results [see @hastieElementsStatisticalLearning2009].

<span id = "reftibshirani">To understand how regularized regression introduces bias, consider a comparison between ordinary least squares regression and LASSO regression [for a more elaborate introduction, see @tibshiraniRegressionShrinkageSelection2011].</span>
Ordinary least squares regression (OLS) estimates model parameters by minimizing the Residual Sum of Squares (RSS) of the outcome variable, given by the formula below.
In this equation, $n$ is the number of participants; $y_i$ is the outcome variable, and $x$ is one of $p$ predictor variables.
\begin{align}
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
\end{align}
The resulting parameter estimates perfectly describe linear relations in the present data set, but generalize less well to new data.
Regularized regression biases parameter estimates towards zero by adding a penalty term to the RSS.
Most common is the LASSO penalty, which consists of the sum of the absolute regression coefficients, or the L1 norm [@hastieElementsStatisticalLearning2009].
As the LASSO penalty is a function of the regression coefficients, it increases when they get bigger.
This incentivizes the optimizer to keep the regression coefficients as small as possible.
The amount of regularization can be controlled by multiplying the penalty by a tuning parameter, $\lambda$.
If $\lambda$ is zero, the shrinkage penalty has no impact.
As $\lambda$ increases,
all coefficients shrink towards zero, ultimately producing the null model.
Cross-validation is often used to find the optimal value for the penalty parameter $\lambda$.
The LASSO penalized residual sum of squares is given by:
\begin{align}
PRSS= RSS + \lambda \sum_{j=1}^{p}|\beta_{j}| (\#eq:prss)
\end{align}
<span id = "whylasso">Note that many other regularizing penalties exist.
This introduction focuses on the LASSO penalty because it is most ubiquitous, easy to understand,
and has an analogue in Bayesian estimation,
as explained in the next section.</span>

@@

A regularizing prior distribution shrinks small coefficients towards zero by assigning high probability mass to near-zero values.
There are many different regularizing prior distributions,
some of which are analogous to specific frequentist methods [@vanErpOberskiMulder2019].
For example, a double exponential prior (hereafter: LASSO prior) results in posterior distributions whose modes are identical to the estimates from LASSO-penalized regression [@ParkCasella2008].
Both the frequentist LASSO penalty and the Bayesian LASSO prior have a tuning parameter $\lambda$ that controls the amount of regularization.
In frequentist LASSO, its value is usually chosen via cross-validation [@hastieElementsStatisticalLearning2009]. 
In the Bayesian approach, by contrast,
a diffuse hyperprior can be used to optimize its value during model estimation [@ParkCasella2008].

@@

The BRMA method introduced here offers both LASSO and regularized horseshoe priors.
The LASSO prior is given by:
\begin{align}
\beta_{j} &\sim \text{DE}(0, \frac{s}{\lambda})\\ (\#eq:lasso)
\frac{1}{\lambda} &\sim \chi^2(0, \nu_1)
\end{align}
where DE denotes the double exponential distribution centered around zero,
with a scale determined by a global scale parameter $s$ that is multiplied by the inverse of tuning parameter $\lambda$.
Increasing the scale parameter extends the prior to cover more extreme values.
The inverse tuning parameter is estimated from the data 
by assigning it a diffuse hyperprior:
a $\chi^2$ prior distribution with mean zero and degrees of freedom $\nu_1$ [@ParkCasella2008].
Increasing the degrees of freedom assigns greater probability mass to extreme values, resulting in less regularization.
The present study used default values for the prior parameters,
$s = 1$, $\nu_1 = 1$, were chosen as a diffuse hyperprior [@ParkCasella2008].

@@

We performed a simulation study to validate the BRMA algorithm.
As a benchmark for comparison, we used random effects meta-regression with restricted maximum likelihood estimation [RMA, @viechtbauerConductingMetaanalysesMetafor2010],
which is the current state-of-the-art in the field.
We evaluated the algorithms' predictive performance in new data, 
ability to perform variable selection,
and ability to recover population parameters.
Our research questions are whether BRMA offers a performance advantage over RMA in terms of any of these indicators,
and which prior (LASSO versus regularized horseshoe) is to be preferred.
<span id = "defaultvalues">For both Bayesian priors, we used default values proposed in prior literature, see Table \@ref(tab:tabprior).
Default values for the LASSO prior were based on @ParkCasella2008,
and default values for the regularized horseshoe prior were based on @PiironenVehtari2017a.</span>
All analysis code is available in a version-controlled repository at <https://github.com/cjvanlissa/pema>.

@@

This example uses the `bonapersona` data, which were included in the `pema` package with permission of the author [@bonapersonaBehavioralPhenotypeEarly2019].
<span id = "sample400">This meta-analysis of over 400 experiments investigated the effects of early life adversity on cognitive performance in rodents.
Note that the sample is much larger than the maximum used to validate BRMA in our simulation study.
As larger samples provide greater statistical power,
it should also be valid for this sample.</span>
For illustrative purposes, we use a smaller subset of the more than 30 moderators.
See the `pema` package documentation (help and vignettes) for further examples.
<!-- We focus specifically on how to choose the different hyperparameters of the regularized horseshoe prior. -->

```{r exdata, eval = FALSE, echo = TRUE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
```
```{r, eval = run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
# Select data to analyze
df <- bonapersona[ , c("yi", "vi", "mTimeLength",
                       "year", "model", "ageWeek")] 
# Multiple imputation for missing values
df <- mice(df)
saveRDS(df, "df_imputed.RData")
```
```{r, eval = !run_everything, echo = FALSE}
# Load relevant packages
library(pema)
library(mice)
library(rstan)
library(ggplot2)
df <- readRDS("df_imputed.RData")
```
<!-- Our simulation study shows good performance with default hyperparameters. -->
<!-- However, experienced users may want to customize the prior. -->
<!-- Visualizing the prior can be helpful in this process. -->
<!-- This is accomplished using the interactive application visualization application available through `shiny_prior()`. -->
<!-- The user can plot the prior distributions resulting from different sets of hyperparameters and compare them. -->
<!-- Increasing the values of the scale parameters (`scale_global` and `hs_scale_slab`) results in a more spread out prior, which applies less regularization. -->
<!-- Increasing the degrees of freedom (`df_global` and `df_slab`) results in thinner tails, which applies more regularization. -->
<!-- One advantage of the regularized horseshoe prior is its ability to define the prior in terms of the expected number of nonzero coefficients, using the argument `relevant_pars`. -->
<!-- This argument in turn influences the other hyperparameters. -->
<!-- We can also plot the implied prior on the effective number of nonzero coefficients given specific hyperparameters. -->
<!-- In this application, we have a total of 7 possible moderators, including the dummy variables for the categorical moderator "model". We can visualize the implied prior on the number of nonzero coefficients for different values of `relevant_pars`: p0 = 1; p0 = 3; p0 = 6; or p0 = NA, in which case no prior information is given and `scale_global`= 1. -->

@@

```{r, echo = FALSE, eval = FALSE}
summary(fit)
```
```{r tabsumex, eval = TRUE, echo = FALSE, results = "asis"}
tab <- summary(fit)$coefficients[1:8, c("mean", "sd", "2.5%", "50%", "97.5%", "n_eff", "Rhat")]
papaja::apa_table(tab, caption = "Summary of model parameters for the applied example.")
```
<!-- , see Figure \@ref(fig:figexres) -->
```{r exres, eval = FALSE, echo = FALSE}
fit_stan <- as.stan(fit)
plot(fit_stan, pars = rownames(coef(fit))[2:8])
```
```{r, eval = run_everything, echo = FALSE}
fit_stan <- as.stan(fit)
p <- plot(fit_stan, pars = rownames(coef(fit))[2:8])
ggsave("plot_es.png", p, device = "png", width = 4, height = 3, dpi = 300, scale = 1.5)
```
```{r figexres, eval = FALSE, echo = FALSE, fig.cap="Visual presentation of model parameters for the applied example."}
knitr::include_graphics("plot_es.png")
```
<!-- Note that there are many auxiliary parameters used for the regularized horseshoe prior. No information on these parameters is returned in the `brma` fitobject, however, once the fitobject is transformed into a Stan object for plotting, these parameters are included. -->

@@

<span id = "sensitivityanalysis">All simulations were conducted with default settings for the model's prior distributions, based on prior research [@ParkCasella2008; @PiironenVehtari2017b].
Our results suggest that these defaults are suitable for a wide range of situations, including when model assumptions are violated.
However, bear in mind that model parameters are influenced by the prior distribution.
It is good practice to perform a sensitivity analysis to determine how sensitive the model results and inferences are to different prior specifications. 
Performing sensitivity analyses is particularly important when the sample is small,
as in this case, the prior is more influential.</span>

@@

<span id = "metareglimits">Before conducting meta-regression,
researchers should be aware of its limitations [see @thompsonHowShouldMetaregression2002].
These can be subdivided into four categories: 1) the curse of dimensionality and its corrolary implications for multicollinearity; 2) the ecological fallacy; 3) limited information on moderator variables, including missing data and restrictions of range.
BRMA seeks to address the first of these limitations,
because the problems that arise from meta-analyzing small and heterogeneous bodies of literature are so ubiquitous that they have been referred to as the primary pitfall in meta-regression [@thompsonHowShouldMetaregression2002].
Nonetheless, all applicable limitations should be acknowledged in the resulting publication.</span>
